---
title: "Sample_Project"
author: "Tuffy Licciardi Issa"
date: "2025-10-25"
output:
  pdf_document:
    latex_engine: xelatex
  html_document:
    number_sections: true
    theme: cerulean
    highlight: tango
    code_folding: show
    df_print: paged
  word_document: default
lang: "pt-BR"
---


> This document presents a minimal and reproducible example of my empirical research workflow.
  
> Full code and additional projects are available at:  **[github.com/tuffyli](https://github.com/tuffyli)**

```{r setup, message=FALSE, warning=FALSE}
set.seed(123)

# Load packages
library(dplyr);library(here);library(tidyverse);library(stargazer);library(MatchIt);library(broom);library(kableExtra);library(cobalt); library(cowplot);library(fixest)

```

## Objective

The objective of this sample is to illustrate an empirical workflow using Brazilian household survey data. I analyze the expansion of alimony rights to women in stable unions and its potential effects on educational attainment among young women.

This policy reform extended alimony eligibility to women in non-civil unions (such as consensual or religious-only unions). The analysis investigates whether this change affected the educational decisions of young women who became newly entitled to this right.

The empirical framework is inspired by Rangel (2006). The main contribution of this exercise is the implementation of an additional Propensity Score Matching (PSM) approach to improve comparability between treated and control individuals.

## Data

The analysis uses microdata from Brazil’s PNAD household survey.

The sample includes women aged 15–24 observed between 1992 and 1995.

Treated individuals are women in non-civil unions (consensual or religious-only) who are household heads or spouses. The control group consists of otherwise similar women in legally recognized civil unions (civil or civil + religious).

Due to the large size of the original PNAD dataset, the full dataset cannot be distributed. The repository therefore includes a filtered dataset containing only the variables required for the empirical analysis. At **[repository](https://github.com/Tuffyli/Personal_Projects/tree/main/PNAD-exercise)**, I include supplementary code detailing the selection of key variables.  

Key variables include:

- age  
- race/color  
- education  
- household per capita income  
- number of children  
- labor market participation  
- state of residence

```{r data, message=FALSE, warning=FALSE}
# ------------------- #
# 1. Data Extraction ----

library(here)

data <- readRDS(here("Ava_Pol","data", "final_filtered_9295.rds"))

# 1.1 Summary ----
# In the data manipulation code I created a summary table. Here I present it
summary_tbl <- readRDS(here("Ava_Pol", "data", "summary.rds"))
print(summary_tbl)
```


# Propensity Score Matching

## Propensity Score Matching

To improve comparability between treated and control individuals, I estimate propensity scores using observable characteristics including age, race, geographic location, household income, and number of children.

Matching is implemented using nearest-neighbour matching with replacement and a caliper of 0.05.

Balance diagnostics are evaluated using standardized mean differences and visualized through love plots. This step allows us to assess whether matching successfully reduces covariate imbalance between treated and control groups before estimating treatment effects.

```{r PSM, warning=FALSE, fig.width=6, fig.height=6, fig.align='center'}
# -------------------------------------------------------------------
# 2. PSM balance
# ------------------------------------------------------------------
plots <- list()

for (year in c(1992, 1993, 1995)) {
  df_psm <- data %>% 
    filter(ano == year) %>% 
    select(
      treatment,                  #Treatment indicator (1 = treated, 0 = control)
      uf,                         #State of residence
      age,                        #Age
      cor,                        #Color/Race
      anos_estudo,                #Years of education
      peso_pessoa,                #Observation weight
      ultima_serie_concluida,     #Last completed grade
      tipo_curso_frequenta,       #Type of course attended
      grupo_cbo,                  #Occupational group
      trabalhou_ultimo_ano,       #Worked in the last 12 months
      renda_dom_per_capita,       #Household per capita income
      dummy_filhos_homens_dom,    #Children Dummy for living with Father
      dummy_filhos_mulheres_dom,  #Children Dummy for living with Mother
      pensao_dummy,               #Pension dummy
      total_filhos                #Total children
    ) %>% 
    mutate(uf = as.factor(uf)) %>% 
    filter(
      !is.na(treatment), !is.na(uf), !is.na(cor), !is.na(peso_pessoa),
      !is.na(dummy_filhos_homens_dom), !is.na(dummy_filhos_mulheres_dom),
      !is.na(renda_dom_per_capita), !is.na(total_filhos)
    )

  match_model <- matchit(
    treatment ~ age + cor + uf + dummy_filhos_homens_dom +
      dummy_filhos_mulheres_dom + renda_dom_per_capita + total_filhos,
    data = df_psm,
    s.weights = df_psm$peso_pessoa,
    caliper = 0.05,
    method = "nearest",
    replace = TRUE
  )

  # Show the match summary counts in the document
  sum <- summary(match_model)
  latex_df <- as.data.frame(sum$nn)
  latex_table <- kable(latex_df,
                       format = "latex",
                       booktabs = TRUE,
                       caption = "Estatísticas descritivas ponderadas",
                       align = "lccc")

  #The files are availabre in the repository
  writeLines(latex_table, 
             paste0("C:/Users/tuffy/Documents/Trabalhos/", year,"_match_summary.tex"))


  # Love plot (INLINE)
  custom_labels <- c(
    distance_weighted         = "Distance (weighted)",
    uf                        = "UF",
    cor                       = "Race/Color",
    dummy_filhos_homens_dom   = "Male child",
    dummy_filhos_mulheres_dom = "Female child",
    age                       = "Age",
    renda_dom_per_capita      = "H. per-capita income",
    total_filhos              = "Total children"
  )

  love <- love.plot(
    match_model,
    stats         = "mean.diffs",
    abs           = TRUE,
    threshold     = 0.10,
    colors        = c("black", "red"),
    line          = TRUE,
    var.names     = custom_labels,
    drop.distance = TRUE
  ) +
    labs(
      title = paste0("Covs. Adjustment: ",year)
    ) +
   theme_minimal(base_size = 5) +
    theme(plot.title = element_text(face = "bold"),
                   legend.position = "bottom")
  #Grouping the plots
  plots[[as.character(year)]] <- love


  matched_data <- match.data(match_model) %>%
    mutate(ano = year)

  assign(paste0("df_matched_", year), matched_data)
  rm(year, custom_labels, sum, ess_table, latex_df, latex_table, love)

}

top    <- cowplot::plot_grid(plots[["1992"]], plots[["1993"]],
                             ncol = 2, align = "hv", rel_widths = c(1, 1))
bottom <- cowplot::plot_grid(plots[["1995"]],
                             ncol = 2, align = "hv", rel_widths = c(1, 1))  # legend stays here

combined <- cowplot::plot_grid(top, bottom, ncol = 1, rel_heights = c(1, 1.05))
combined



```

```{r clean, message=FALSE, warning=FALSE, include=FALSE}
rm(bottom, top, combined, match_model, plots)
```


## Regression

To estimate treatment effects, I fit progressively richer regression specifications using the matched sample.

The outcome variable is years of education.

Three specifications are estimated:

1. Baseline model without fixed effects.
2. Model including state-by-year fixed effects to control for geographic shocks.
3. Model including fixed effects and additional individual controls.

All regressions are estimated using matching weights and heteroskedasticity-robust standard errors.


```{r}
# ---------------------------------------------------------------------------- #
# 3. Regression -----
# ---------------------------------------------------------------------------- #
# Bind matched samples (1992, 1993, 1995) and prepare identifiers
df_psm <- bind_rows(df_matched_1992, df_matched_1993, df_matched_1995)

df_psm$ano   <- factor(df_psm$ano)        # ensure year is a factor
df_psm$ano   <- relevel(df_psm$ano, ref = "1992")   # reference year = 1992
df_psm$uf_ano <- interaction(df_psm$uf, df_psm$ano, sep = "_")  # UF × year FE key


# --- Model 1: no fixed effects (year-to-year comparison via i())
model_nc <- feols(
  anos_estudo ~ treatment + i(ano, treatment, ref = "1992"),
  data = df_psm, weights = df_psm$weights, vcov = "hetero"
)

# --- Model 2: add UF×year fixed effects (absorbs geography-by-year shocks)
model_fe <- feols(
  anos_estudo ~ treatment + i(ano, treatment, ref = "1992") | uf_ano,
  data = df_psm, weights = df_psm$weights, vcov = "hetero"
)

# --- Model 3: UF×year FE + basic controls
model_cc <- feols(
  anos_estudo ~ treatment + i(ano, treatment, ref = "1992") +
    total_filhos + cor + age | uf_ano,
  data = df_psm, weights = df_psm$weights, vcov = "hetero"
)
```

```{r clean2, message=FALSE, warning=FALSE, include=FALSE}
rm(df_matched_1992, df_matched_1993, df_matched_1995, matched_data)
```


```{r Output, results='asis'}
tab <- fixest::etable(
  model_nc, model_fe, model_cc,            # pass models directly
  drop      = "Constant",
  dict = c(
    "anos_estudo"                    = "Years of education",
    "treatment$"                     = "Treatment (1992)",
    "i\\(ano, treatment.*\\)::1993"  = "Treatment × 1993",
    "i\\(ano, treatment.*\\)::1995"  = "Treatment × 1995",
    "age"                            = "Age",
    "cor"                            = "Race/Color",
    "total_filhos"                   = "Total children",
    "treatment x ano = 1993"         = "Treatment × 1993",
    "treatment x ano = 1995"         = "Treatment × 1995"
  ),
  se.below   = TRUE,
  signif.code = "letters",
  digits     = 3,
  fitstat    = c("n","r2","rmse"),
  title      = "ATT by Year (ref = 1992)",
  notes      = "SEs are heteroskedastic-robust. Weights from matching sample."
  # <- DO NOT set tex=TRUE here (default is HTML/markdown-friendly)
)

knitr::kable(tab, format = "latex", escape = FALSE) |>
  kableExtra::kable_styling(full_width = FALSE)
```

Across all specifications, the estimated treatment effects are not statistically significant.

This suggests that the expansion of alimony eligibility did not produce measurable changes in educational attainment among young women in this sample.

However, the relatively low explanatory power of the models indicates that much of the variation in educational outcomes remains unexplained. These results should therefore be interpreted cautiously.

## Reproducibility

The full empirical project, including data preparation, matching procedures, and regression code, is available in the repository:

github.com/tuffyli

The repository contains the complete scripts and documentation required to replicate the analysis.